prolfqua - An R-package for Proteomics Differential Expression Analysis.

### 2022-05-06 inst/MyArticle/paper.Rmd

We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value[@tidydata2014]. Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table. For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration. For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables. Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base __R__[@RCoreTeam2021] and the tidyverse[@tidyverse2019]. We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)). __R6__ classes are well supported, e.g. by auto-completion features, and easy to use (OOP).

@@

```{r proDAAllBenchmarks, fig.cap="Comparison of fold change estimates produced by proDA and prolfqua simple impute method."}
proda <- allBenchmarks$benchmark_proDA$data()
imp <- allBenchmarks$benchmark_imputation$data()
imp <- allBenchmarks$benchmark_ProtModerated$data()

x <- dplyr::inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".imp",".proda"))
ggplot2::ggplot(x,ggplot2::aes( x = diff.imp, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="lightgray") + ggplot2::geom_vline(xintercept = 0, color="lightgray") + ggplot2::geom_abline(slope = 1) + ggplot2::theme_minimal()

@@

## Package sticker

>>>>

We store all the data needed for analysis in a single data frame in a tidy table, i.e., every column is a variable, every row is an observation, every cell is a single value[@tidydata2014]. Using an __R6__[@R6cite] configuration object, we specify what variable is in which column, making it easy to integrate new inputs in `r BiocStyle::Githubpkg("fgcz/prolfqua")` if provided as a tidy table. For example, to visualize tidy _Spectronaut_, or _Skyline_ outputs, or data in `r BiocStyle::Biocpkg("MSstats")` format, only a few lines of code are needed to update the `r BiocStyle::Githubpkg("fgcz/prolfqua")` configuration. For popular software like _MaxQuant_ or _MSFragger_, which stores the same variable, e.g., intensity, in multiple columns, one for each sample, we implemented methods that transform the data into tidy tables. Relying on the tidy data table enable us to easily interface with many data manipulation, visualization, and modeling methods implemented in base __R__[@RCoreTeam2021] and the tidyverse[@tidyverse2019]. We use __R6__ classes to structure the functionality of the package (see Figure \@ref(fig:LFQData) and Figure \@ref(fig:ContrastUML)). __R6__ classes are well supported, e.g., by command-line completion features, see Figure \@ref{fig:tabCompletion}, and easy to use (OOP).

@@

(ref:proDAAllBenchmarks) Comparison of fold change estimates produced by proDA and using the missing data modelling method implemented in prolfqua.

```{r proDAAllBenchmarks, fig.cap="(ref:proDAAllBenchmarks)"}
proda <- allBenchmarks$benchmark_proDA$data()
imp <- allBenchmarks$benchmark_imputation$data()
imp <- allBenchmarks$benchmark_ProtModerated$data()

x <- dplyr::inner_join(imp, proda, by = c("contrast","protein_Id", "TP", "species"), suffix = c(".lm",".proda"))
ggplot2::ggplot(x,ggplot2::aes( x = diff.lm, y = diff.proda) ) + ggplot2::geom_point() + ggplot2::facet_wrap(~contrast + species) + ggplot2::geom_hline(yintercept = 0, color="lightgray") + ggplot2::geom_vline(xintercept = 0, color="lightgray") + ggplot2::geom_abline(slope = 1) + ggplot2::theme_minimal()

@@

\newpage
## Miscellaneous

(ref:tabCompletion) The screenshot displays the command-line completion (tab completion) of RStudio on the `prolfqua::LFQData` R6 object. In the example, it shows the getter methods of the object.

```{r tabCompletion, echo=FALSE, fig.cap="(ref:tabCompletion)", out.width = '66%'}
knitr::include_graphics("codeSnippet1TabCompletion.png")
```
